The evolution of gene expression in the term placenta of viviparous mammals

### 2015-12-11 manuscript/manuscript.Rnw
make links black

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
%\usepackage[plain]{fancyref}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=reading,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage{cleveref}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","placenta_go_results_mf")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf"
                              ))) {
    load(a)
}
combined.fpkm[,species:=factor(species,levels=names(species.names))]
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
@

@@

<<top.genes.function,echo=FALSE,results="hide">>=
top.genes.table <- function(species.name,n=50) {
    temp <- 
        combined.fpkm[species==species.name &
                           !(!is.na(human_name) & 
                                 human_name %in% housekeeping.genes.superset[,gene_short_name]) &
                               !grepl("((RNA\\d|MT)-|^RP[SL]\\d+$)",human_name) &
                                   !grepl("(^[Mm][Tt]-|^R[pP][lsLS]\\d+[abcdABCD]?[l]?(?:-.+)?$|rRNA$)",gene_short_name)
                      ][order(-mean_fpkm)][,list(name_or_id,
                                                 human_name,mean_fpkm)]
    temp <- temp[!duplicated(name_or_id)][1:n]
    ## remove any genes which have a null name_or_id; they are probably from unassembled contigs
    temp <- temp[!(is.na(name_or_id)|name_or_id==""),]
    setnames(temp,c("Gene name","Human Name","FPKM"))
    if (species.name=="homo sapiens") {
        temp <- temp[,-2,with=FALSE]
    }
    return(temp)
}
gene.expression <- function(species.name,gene,complete=FALSE) {
    fpkm <- combined.fpkm[species==species.name &
                              gene_short_name==gene,
                          mean_fpkm]

@@

\begin{table}
  \tiny
<<top.50.human.expressed.genes,echo=FALSE,results="asis">>=
top.human.genes <- top.genes.table(species.name="homo sapiens")
print(xtable(top.human.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.human.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not
  ribosomal, mitochondrial, or a housekeeping gene in \textit{Homo
    sapiens}}
\label{tab:top50_human_placenta}
\end{table}

@@

\begin{table}
  \tiny
<<top.50.bonobo.expressed.genes,echo=FALSE,results="asis">>=
top.bonobo.genes <- top.genes.table(species.name="pan paniscus")
print(xtable(top.bonobo.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.bonobo.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Pan paniscus}}
\label{tab:top50_bonobo_placenta}
\end{table}

@@

\begin{table}
  \tiny
<<top.50.spider.monkey.expressed.genes,echo=FALSE,results="asis">>=
top.spider.monkey.genes <- top.genes.table(species.name="ateles fusciceps")
print(xtable(top.spider.monkey.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.spider.monkey.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Ateles fusciceps}}
\label{tab:top50_spider_monkey_placenta}
\end{table}

@@

\begin{table}
  \tiny
<<top.50.dog.expressed.genes,echo=FALSE,results="asis">>=
top.dog.genes <- top.genes.table(species.name="canis familiaris")
print(xtable(top.dog.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.dog.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Canis familiaris}}
\label{tab:top50_dog_placenta}
\end{table}

The top 50 expressed genes as measured by \ac{FPKM} in the term placenta of
the dog are shown in Table 4 with a list of all genes in Supplemental
Table X. The gene with the highest expression (\ac{FPKM}=48399.9) in the
dog placenta is the hormone relaxin (RLN1). Other highly expressed
genes among the top 50 (Table 4) include genes involved in the
cytoskeleton (TMSB10, TMSB4X), anticoagulation (TFPI2, PLA2G7), innate
immunity (LCN2, GM2A), lipid metabolism (APOE), inhibition of matrix
metalloproteinases (TIMP2, TIMP3), and growth (PGF, EGFR, HTRA1). The
gene LGALS1, which shows evidence of immunosupression during pregnancy
(Than et al., 2008), shows high expression in dog (\ac{FPKM}= 1405.02). In
addition, the gene SSLP1, a gene expressed in the placentas of cow and
sheep but not present in primates (Ushizawa et al., 2009), is also
expressed at high levels (\ac{FPKM}= 1853.98). There is no placental
expression of the estrogen synthase gene (CYP19A1) in dogs, however
expression of HSD3B1 is relatively high (\ac{FPKM}=157.622).

@@

\begin{table}
  \tiny
<<top.50.cow.expressed.genes,echo=FALSE,results="asis">>=
top.cow.genes <- top.genes.table(species.name="bos taurus")
print(xtable(top.cow.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.cow.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Bos taurus}}
\label{tab:top50_cow_placenta}
\end{table}

@@

\begin{table}
\tiny
<<top.50.elephant.expressed.genes,echo=FALSE,results="asis">>=
top.elephant.genes <- top.genes.table(species.name="loxodonta africana")
print(xtable(top.elephant.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.elephant.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Loxodonta africana}}
\label{tab:top50_elphant_placenta}
\end{table}

@@

\begin{table}
  \tiny
<<top.50.opossum.expressed.genes,echo=FALSE,results="asis">>=
top.opossum.genes <- top.genes.table(species.name="monodelphis domestica")
print(xtable(top.opossum.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.opossum.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Monodelphis domestica}}
\label{tab:top50_opossum_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.mouse.expressed.genes,echo=FALSE,results="asis">>=
top.mouse.genes <- top.genes.table(species.name="mus musculus")
print(xtable(top.mouse.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.mouse.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Mus musculus}}
\label{tab:top50_mouse_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.n.galili.expressed.genes,echo=FALSE,results="asis">>=
top.galili.genes <- top.genes.table(species.name="nannospalax galili")
print(xtable(top.galili.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.galili.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Nannospalax galili}}
\label{tab:top50_n_galili_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.s.carmeli.expressed.genes,echo=FALSE,results="asis">>=
top.carmeli.genes <- top.genes.table(species.name="spalax carmeli")
print(xtable(top.carmeli.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.carmeli.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Spalax carmeli}}
\label{tab:top50_s_carmeli_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.pig.expressed.genes,echo=FALSE,results="asis">>=
top.pig.genes <- top.genes.table(species.name="sus scrofa")
print(xtable(top.pig.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.pig.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Sus scrofa}}
\label{tab:top50_pig_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.horse.expressed.genes,echo=FALSE,results="asis">>=
top.horse.genes <- top.genes.table(species.name="equus caballus")
print(xtable(top.horse.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.horse.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Equus caballus}}
\label{tab:top50_horse_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.sheep.expressed.genes,echo=FALSE,results="asis">>=
top.sheep.genes <- top.genes.table(species.name="ovis aries")
print(xtable(top.sheep.genes),
      booktabs=TRUE,
      floating=FALSE
      )
setkey(top.sheep.genes,"Gene name")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Ovis aries}}
\label{tab:top50_sheep_placenta}
\end{table}

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
%\usepackage[plain]{fancyref}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=reading,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage{cleveref}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","placenta_go_results_mf","hugo_nomenclature")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf",
                              "hugo_nomenclature"
                              ))) {
    load(a)
}
combined.fpkm[,species:=factor(species,levels=names(species.names))]
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
setkey(hugo.nomenclature,"Approved Symbol")
@

@@

<<top.genes.function,echo=FALSE,results="hide">>=
top.genes.table <- function(species.name,n=50) {
    temp <- 
        combined.fpkm[species==species.name &
                           !(!is.na(human_name) & 
                                 human_name %in% housekeeping.genes.superset[,gene_short_name]) &
                               !grepl("((RNA\\d|MT)-|^RP[SL]\\d+$)",human_name) &
                                   !grepl("(^[Mm][Tt]-|^R[pP][lsLS]\\d+[abcdABCD]?[l]?(?:-.+)?$|rRNA$)",gene_short_name)
                      ][order(-mean_fpkm)][,list(name_or_id,
                                                 human_name,mean_fpkm)]
    ## remove any genes which have a null name_or_id; they are probably from unassembled contigs
    temp <- temp[!(is.na(name_or_id)|name_or_id==""),]
    temp <- temp[!duplicated(name_or_id)][1:n]
    setkey(temp,"human_name")
    temp <- 
        hugo.nomenclature[temp][order(-mean_fpkm),
                                list(name_or_id,`Approved Symbol`,`Approved Name`,mean_fpkm)]
    temp[!is.na(`Approved Name`),
         `Approved Name`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{",xtable.sanitize.text(`Approved Name`),"}")]
    temp[!is.na(`Approved Symbol`),
         `Approved Symbol`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{",xtable.sanitize.text(`Approved Symbol`),"}")]
    setnames(temp,c("Gene symbol","Human Symbol","Human Name","FPKM"))
    if (species.name=="homo sapiens") {
        temp <- temp[,-2,with=FALSE]
    }
    return(temp)
}
genes.table.sanitize <- function(str){
    return(str)
}
xtable.sanitize.text <- function(str){
    result <- str
    result <- gsub("\\\\", "SANITIZE.BACKSLASH", result)
    result <- gsub("$", "\\$", result, fixed = TRUE)
    result <- gsub(">", "$>$", result, fixed = TRUE)
    result <- gsub("<", "$<$", result, fixed = TRUE)
    result <- gsub("|", "$|$", result, fixed = TRUE)
    result <- gsub("{", "\\{", result, fixed = TRUE)
    result <- gsub("}", "\\}", result, fixed = TRUE)
    result <- gsub("%", "\\%", result, fixed = TRUE)
    result <- gsub("&", "\\&", result, fixed = TRUE)
    result <- gsub("_", "\\_", result, fixed = TRUE)
    result <- gsub("#", "\\#", result, fixed = TRUE)
    result <- gsub("^", "\\verb|^|", result, fixed = TRUE)
    result <- gsub("~", "\\~{}", result, fixed = TRUE)
    result <- gsub("SANITIZE.BACKSLASH", "$\\backslash$",
                   result, fixed = TRUE)
    return(result)
}
gene.expression <- function(species.name,gene,complete=FALSE) {
    fpkm <- combined.fpkm[species==species.name &
                              gene_short_name==gene,
                          mean_fpkm]

@@

\begin{table}
  \tiny
<<top.50.human.expressed.genes,echo=FALSE,results="asis">>=
top.human.genes <- top.genes.table(species.name="homo sapiens")
print(xtable(top.human.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.human.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not
  ribosomal, mitochondrial, or a housekeeping gene in \textit{Homo
    sapiens}}
\label{tab:top50_human_placenta}
\end{table}

@@

\begin{table}
  \tiny
<<top.50.bonobo.expressed.genes,echo=FALSE,results="asis">>=
top.bonobo.genes <- top.genes.table(species.name="pan paniscus")
print(xtable(top.bonobo.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.bonobo.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Pan paniscus}}
\label{tab:top50_bonobo_placenta}
\end{table}

@@

\begin{table}
  \tiny
<<top.50.spider.monkey.expressed.genes,echo=FALSE,results="asis">>=
top.spider.monkey.genes <- top.genes.table(species.name="ateles fusciceps")
print(xtable(top.spider.monkey.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.spider.monkey.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Ateles fusciceps}}
\label{tab:top50_spider_monkey_placenta}
\end{table}

@@

\begin{table}
  \tiny
<<top.50.dog.expressed.genes,echo=FALSE,results="asis">>=
top.dog.genes <- top.genes.table(species.name="canis familiaris")
print(xtable(top.dog.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.dog.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Canis familiaris}}
\label{tab:top50_dog_placenta}
\end{table}

The top 50 non-housekeeping expressed genes as measured by \ac{FPKM}
in the term placenta of \textit{Canis familiaris} are shown in
\cref{tab:top50_dog_placenta} with a list of all genes in Supplemental
\cref{tab:all_canis_familiaris_placenta}. The gene with the highest
expression %
(\ac{FPKM}=\Sexpr{gene.expression("canis
  familiaris","ENSCAFG00000002115")}) %
is the hormone relaxin (\textit{RLN2}). Other highly expressed genes
among the top 50 (\cref{tab:top50_dog_placenta}) include genes
involved in the cytoskeleton (\textit{TMSB10}, \textit{TMSB4X}),
anticoagulation (\textit{TFPI2}, \textit{PLA2G7}), innate immunity
(\textit{LCN2}, \textit{GM2A}), lipid metabolism (\textit{APOE}),
inhibition of matrix metalloproteinases (\textit{TIMP2},
\textit{TIMP3}), and growth factors (\textit{PGF}, \textit{EGFR},
HTRA1). The gene \textit{LGALS1}, which shows evidence of
immunosupression during pregnancy
\cite{Than.ea2008:Emergenceofhormonaland}, shows high expression in
dog %
(\ac{FPKM}=\Sexpr{gene.expression("canis familiaris","LGALS1")}). %
In addition, the gene SSLP1, a gene expressed in the placentas of cow
and sheep but not present in primates (Ushizawa et al., 2009), is also
expressed at high levels
%
(\ac{FPKM}=\Sexpr{gene.expression("canis familiaris","SSLP1")}). %
There is no placental expression of the estrogen synthase gene
(CYP19A1) in dogs, however expression of HSD3B1 is relatively high
(\ac{FPKM}=\Sexpr{gene.expression("canis familiaris","HSD3B1")}).

@@

\begin{table}
  \tiny
<<top.50.cow.expressed.genes,echo=FALSE,results="asis">>=
top.cow.genes <- top.genes.table(species.name="bos taurus")
print(xtable(top.cow.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.cow.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Bos taurus}}
\label{tab:top50_cow_placenta}
\end{table}

@@

\begin{table}
\tiny
<<top.50.elephant.expressed.genes,echo=FALSE,results="asis">>=
top.elephant.genes <- top.genes.table(species.name="loxodonta africana")
print(xtable(top.elephant.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.elephant.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Loxodonta africana}}
\label{tab:top50_elphant_placenta}
\end{table}

@@

\begin{table}
  \tiny
<<top.50.opossum.expressed.genes,echo=FALSE,results="asis">>=
top.opossum.genes <- top.genes.table(species.name="monodelphis domestica")
print(xtable(top.opossum.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.opossum.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Monodelphis domestica}}
\label{tab:top50_opossum_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.mouse.expressed.genes,echo=FALSE,results="asis">>=
top.mouse.genes <- top.genes.table(species.name="mus musculus")
print(xtable(top.mouse.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.mouse.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Mus musculus}}
\label{tab:top50_mouse_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.n.galili.expressed.genes,echo=FALSE,results="asis">>=
top.galili.genes <- top.genes.table(species.name="nannospalax galili")
print(xtable(top.galili.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.galili.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Nannospalax galili}}
\label{tab:top50_n_galili_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.s.carmeli.expressed.genes,echo=FALSE,results="asis">>=
top.carmeli.genes <- top.genes.table(species.name="spalax carmeli")
print(xtable(top.carmeli.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.carmeli.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Spalax carmeli}}
\label{tab:top50_s_carmeli_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.pig.expressed.genes,echo=FALSE,results="asis">>=
top.pig.genes <- top.genes.table(species.name="sus scrofa")
print(xtable(top.pig.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.pig.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Sus scrofa}}
\label{tab:top50_pig_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.horse.expressed.genes,echo=FALSE,results="asis">>=
top.horse.genes <- top.genes.table(species.name="equus caballus")
print(xtable(top.horse.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.horse.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Equus caballus}}
\label{tab:top50_horse_placenta}
\end{table}

@@

\begin{table}[h]
  \tiny
<<top.50.sheep.expressed.genes,echo=FALSE,results="asis">>=
top.sheep.genes <- top.genes.table(species.name="ovis aries")
print(xtable(top.sheep.genes),
      booktabs=TRUE,
      floating=FALSE,
      sanitize.text.function=genes.table.sanitize,include.rownames=FALSE)
setkey(top.sheep.genes,"Gene symbol")
@ 
\caption{Expression in \ac{FPKM} of top 50 genes which are not ribosomal, 
mitochondrial, or whose human ortholog is a housekeeping gene in
  \textit{Ovis aries}}
\label{tab:top50_sheep_placenta}
\end{table}
